---
title: "2018 early voting and absentee"
output: html_document
date: "2022-11-08"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


#Loading the packages & dataframes we need. 
```{r}
library(readr)
library(tidyverse)
absentee_gender_18 <- read_csv("absentee_gender_18.csv") 
absentee_ages_18 <- read_csv("absentee_ages_18.csv")
ev_gender_18 <- read_csv("ev_gender_18.csv") 
ev_ages_18 <- read_csv("ev_ages_18.csv")

absentee_gender_18 <- absentee_gender_18 %>% filter(!is.na(County))
```


#Combining early voting & absentee
```{r}
gender_and_county_18 <- union_all(absentee_gender_18,ev_gender_18)
age_and_county_18 <- union_all(absentee_ages_18,ev_ages_18)

```

#Making a dataframe that shows us how many early voting and absentee ballots were cast in each county, split by gender. These are raw numbers, not adjusted for registration totals. 
```{r}
gender_and_county_18 <- tibble::rowid_to_column(gender_and_county_18,"ID")
gender_and_county_18 <- gender_and_county_18 %>% group_by(County,`Voter Gender`) %>% summarise(ID = n()) %>% pivot_wider(names_from = `Voter Gender`, values_from = ID) 

#Warning. Text errors that will be cleaned in excel. Such as saying N instead of NA. OR "m" instead of "M" for male.
write.csv(absentee_and_early_voting_split_by_gender_and_county_18,"absentee_and_early_voting_split_by_gender_and_county_18.csv")
```

#Making a dataframe that shows us how many early voting and absentee ballots were cast in each county, split by age brackets These are raw numbers, not adjusted for registration totals. 
```{r}
#First, we make individual dataframes for each age bracket. 
a <- age_and_county_18 %>% filter(`Voter Age`>= 18,`Voter Age`<=24.999) %>% mutate(age_bracket = "18 to 24")
b <- age_and_county_18%>% filter(`Voter Age`>= 25,`Voter Age`<=34.999) %>% mutate(age_bracket = "25 to 34")
c <- age_and_county_18%>% filter(`Voter Age`>= 35,`Voter Age`<=44.999) %>% mutate(age_bracket = "35 to 44")
d <- age_and_county_18 %>% filter(`Voter Age`>= 45,`Voter Age`<=54.999) %>% mutate(age_bracket = "45 to 54")
e <- age_and_county_18 %>% filter(`Voter Age`>= 55,`Voter Age`<=64.9999) %>% mutate(age_bracket = "55 to 64")
f <- age_and_county_18%>% filter(`Voter Age`>= 65,`Voter Age`<=74.9999) %>% mutate(age_bracket = "65 to 74")
g <- age_and_county_18 %>% filter(`Voter Age`>= 75) %>% mutate(age_bracket = "75 and up ")

#And now we recombine the dataframe, with every voter classified into an age bracket, according to age.
age_and_county_18_grouped <- union_all(a,b)
age_and_county_18_grouped  <- union_all(age_and_county_18_grouped,c)
age_and_county_18_grouped  <- union_all(age_and_county_18_grouped ,d)
age_and_county_18_grouped  <- union_all(age_and_county_18_grouped ,e)
age_and_county_18_grouped  <- union_all(age_and_county_18_grouped ,f)
age_and_county_18_grouped  <- union_all(age_and_county_18_grouped ,g)

```

#Now let's make a dataframe that showsus how many early voting and absentee ballots were cast in each county, split by age bracket. These are raw numbers, not adjusted for registration totals. 
```{r}

absentee_and_early_voting_split_by_age_and_county_18 <-age_and_county_18_grouped %>% group_by(County,age_bracket) %>% summarise(`Voter Age` = n()) %>% pivot_wider(names_from = age_bracket, values_from = `Voter Age`) 

write.csv(absentee_and_early_voting_split_by_age_and_county_18,"absentee_and_early_voting_split_by_age_and_county_18.csv")
```




