---
title: "voter registration 2018 and 2022"
output: html_document
date: "2022-11-10"
---
#Load in the dataframes & rename them  
```{r}
library(tidyverse)
library(lubridate)
library(readr)
library(stringr)

voter_registration_2022 <-StateVR101822 <- read_delim("StateVR101822.txt", 
    delim = "|", escape_double = FALSE, trim_ws = TRUE)

voter_registration_2018 <- read_delim("tester.txt", delim = "|", 
    escape_double = FALSE, trim_ws = TRUE)
View(tester)

```

#Cleaning up the birthday column because good lord 
```{r}

#Pulling out the year and month of birth
voter_registration_2022 <-voter_registration_2022 %>% mutate(year_of_birth = str_sub(voter_registration_2022$`DATE OF BIRTH`,-12,-9),extra_months = str_sub(voter_registration_2022$`DATE OF BIRTH`,1,1))

#Making sure the computer knows they're numbers,so we can do some math 
voter_registration_2022$extra_months <-as.numeric(voter_registration_2022$extra_months)
voter_registration_2022$year_of_birth <-as.numeric(voter_registration_2022$year_of_birth)

#Doing math, so we don't count someone in the wrong age bracket by a couple of months. 
voter_registration_2022 <- voter_registration_2022 %>% mutate(extra_months_converted = extra_months/12)

voter_registration_2022 <- voter_registration_2022 %>% mutate(year_and_month_converted = year_of_birth+extra_months_converted)

#Rigamoroling this backwards to find age
voter_registration_2022 <- voter_registration_2022 %>% mutate(age = 2022.91 - year_and_month_converted)


```

#Making a dataframe that shows us how many voters are registered each county, split by age brackets These are raw numbers, not adjusted for registration totals. 
```{r}
#First, we make individual dataframes for each age bracket. 
a <- voter_registration_2022 %>% filter(age>= 18,`age`<=24.999) %>% mutate(age_bracket = "18 to 24")
b <- voter_registration_2022 %>% filter(`age`>= 25,`age`<=34.999) %>% mutate(age_bracket = "25 to 34")
c <- voter_registration_2022 %>% filter(`age`>= 35,`age`<=44.999) %>% mutate(age_bracket = "35 to 44")
d <- voter_registration_2022  %>% filter(`age`>= 45,`age`<=54.999) %>% mutate(age_bracket = "45 to 54")
e <- voter_registration_2022 %>% filter(`age`>= 55,`age`<=64.9999) %>% mutate(age_bracket = "55 to 64")
f <- voter_registration_2022 %>% filter(`age`>= 65,`age`<=74.9999) %>% mutate(age_bracket = "65 to 74")
g <- voter_registration_2022 %>% filter(`age`>= 75) %>% mutate(age_bracket = "75 and up ")

#And now we recombine the dataframe, with every voter classified into an age bracket, according to age.
voter_registration_2022_turnout_grouped <- union_all(a,b)
voter_registration_2022_turnout_grouped  <- union_all(voter_registration_2022_turnout_grouped,c)
voter_registration_2022_turnout_grouped  <- union_all(voter_registration_2022_turnout_grouped ,d)
voter_registration_2022_turnout_grouped  <- union_all(voter_registration_2022_turnout_grouped ,e)
voter_registration_2022_turnout_grouped  <- union_all(voter_registration_2022_turnout_grouped ,f)
voter_registration_2022_turnout_grouped  <- union_all(voter_registration_2022_turnout_grouped ,g)

```


#Now let's make a dataframe that showsus how many folks were registered in each county, split by age bracket. 
```{r}
voter_registration_2022_turnout_grouped <-voter_registration_2022_turnout_grouped %>% group_by(COUNTY_NAME,age_bracket) %>% summarise(`DATE OF BIRTH` = n()) %>% pivot_wider(names_from = age_bracket, values_from = `DATE OF BIRTH`) 


write.csv(voter_registration_2022_turnout_grouped ,"voter_registration_split_by_age_and_county_22.csv")
```

#Making a dataframe that shows us how many folks were registered in each county, split by gender.
```{r}
voter_registration_2022 <- tibble::rowid_to_column(voter_registration_2022,"ID")
voter_registration_2022 %>% group_by(COUNTY_NAME,SEX) %>% summarise(ID = n()) %>% pivot_wider(names_from = SEX, values_from = ID) %>% mutate(Unsure = Unknown+`NA`) 


#write.csv(voter_registration_split_by_gender_and_county_22,"voter_registration_split_by_gender_and_county_22.csv")

```

#Making a dataframe showing how many voters were registered in each county
```{r}
total_turnout_by_county_2022 <- voter_registration_2022 %>% group_by(COUNTY_NAME) %>% summarise(ID = n()) 


write.csv(total_turnout_by_county_2022,"total_turnout_by_county.csv")

```




